%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{SPMD Execution Model}

\begin{itemize}
    \item Data-parallel
    \begin{itemize}
        \item Work needs to be divided
    \end{itemize}
    
    \item Single program, scalar with implicit SIMD execution
    
    \item Multiple instances running in parallel
    \begin{itemize}
        \item Each instance working on different data
        \item On GPU, SIMD execution in lockstep
        \item On CPU, sequential execution within a core (naive approach)
    \end{itemize}
\end{itemize}

\begin{itemize}
    \item Simplistic example:
    \begin{itemize}
        \item Massive Online Course
        \item Compute the overall grade of millions of students
        \item Overall student's grade is the weighted average of several grades
    \end{itemize}
\end{itemize}

\end{frame}

%% Kernel function has no return value, takes in buffers (arrays)
%% Vectorization does not change the signature for kernels
%% Can also vectorize "normal" functions, but this results in signature changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Division of Work}

\begin{itemize}
    \item Work-item:
    \begin{itemize}
        \item Unit of work
        \item One instance of a program
    \end{itemize}
    \item Work-items executed in parallel by Execution Units
    \item Dimensions
    \begin{itemize}
        \item 1D (array shape)
        \item 2D (grid shape)
        \item ...
    \end{itemize}
\end{itemize}

[graph]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Single Program}

\begin{itemize}
    \item Kernel function
    \begin{itemize}
        \item Entry point for the computation
    \end{itemize}
    \item Executed once per work-item
    \begin{itemize}
        \item As if there was a loop around it
        \item Access to the iteration counter using \texttt{get\_global\_id(0)}
    \end{itemize}
\end{itemize}

[code]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Vertical Vectorization}

\begin{itemize}
    \item Within a single work-item
    \begin{itemize}
        \item e.g. loops within a kernel
        \item In our example, the weighted average computation
        \end{itemize}
    \item Using the LLVM Loop Vectorizer
    \item However, not all kernels contain loops
\end{itemize}

[graph of work-items with vertical loops]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Horizontal Vectorization}

\begin{itemize}
    \item Many executions units each executing one instance of a single program 
    \begin{itemize}
        \item Fine on GPU (many EUs)
        \item Not so much on CPU (very few EUs)
        \item CPU has to execute many work-items sequentially
        \end{itemize}
    
    \item Across work-items
    \begin{itemize}
        \item Compute multiple work-items at the same time
        \item Take advantage of the execution model
    \end{itemize}
\end{itemize}

[graph of work-items with horizontal arrows]

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Glossary}

\begin{itemize}
    \item Work
    \begin{itemize}
        \item Work-item: unit of work to execute in parallel.
        \item Instance: State of one work-item.
        \item (SIMD) Lane: Execution of one instance, after vectorization.
        \end{itemize}
        
    \item Data
    \begin{itemize}
        \item Packet: Contains several values/instructions, one per SIMD lane. Corresponds to one value/instruction in the original kernel.
        \item \uniform{Uniform}: Packet where values are identical for all lanes.
        \item \varying{Varying}: Packet where values are not identical for all lanes.
        \end{itemize}

    \item Control Flow
    \begin{itemize}
        \item \uniform{Uniform}: Branch taken by all lanes.
        \item \varying{Divergent}: Branch taken by some lanes. Requires special handling (No SIMD branching)
    \end{itemize}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Vectorizer Comparison}

\begin{itemize}
    \item Loop Vectorizer
    \begin{itemize}
        \item Can be used for both vertical and horizontal vectorization
        \item Has to enforce dependencies between loop iterations
        \item Execution order is not specified, this is not needed
        \item Nested control-flow?
    \end{itemize}

    \item SLP Vectorizer
    \begin{itemize}
        \item Finds groups of similar scalar instructions (same opcode)
        \item Vertical vectorization
        \item Not all kernels contain this kind of code
    \end{itemize}

    \item SPMD Vectorizer
    \begin{itemize}
        \item Only supports horizontal vectorization
        \item Nested control-flow
        \item Not limited to a certain 'style' of code
    \end{itemize}
\end{itemize}

\end{frame}
